Malletwand: the Pendulum as a Handheld Interface to Musical Timing

### 2024-02-08 paper/paper.tex
Paper

% Uncomment only one of the ones below
\usepackage{anonymize} 		   %Uncomment this line to publish
% \usepackage[blind]{anonymize}%Uncomment this line for blind review

@@

\label{key}
\numberofauthors{4}
\author{
\alignauthor
\anonymize{Lyu, S.} \\
  \affaddr{\anonymize{Academy of Arts \& Design, Tsinghua University}}\\
  \affaddr{\anonymize{Beijing, China}}\\
  \email{\anonymize{claire@ayu.land}}
\alignauthor
\anonymize{Li, H.} \\
  \affaddr{\anonymize{Academy of Arts \& Design, Tsinghua University}}\\
  \affaddr{\anonymize{Beijing, China}}\\
  \email{\anonymize{lihanxua22@mails.tsinghua.edu.cn}}
\and
\alignauthor
\anonymize{Wang, R.} \\
  \affaddr{\anonymize{Academy of Arts \& Design, Tsinghua University}}\\
  \affaddr{\anonymize{Beijing, China}}\\
  \email{\anonymize{wangrh22@mails.tsinghua.edu.cn}}
\alignauthor
\anonymize{Mi, H.}\titlenote{\anonymize{Corresponding author.}}\\
  \affaddr{\anonymize{Academy of Arts \& Design, Tsinghua University}}\\
  \affaddr{\anonymize{Beijing, China}}\\
  \email{\anonymize{mhp@tsinghua.edu.cn}}
}
\date{6 February 2024}

@@

The instrument is accompanied by two ``wand'' controllers and can operate in both solo and co-operative mode. Signals received from the controllers drive the passing of musical time in a predetermined musical piece, similar to a musical box. Co-operative mode is entered by two controllers being operated simultaneously, where each controller corresponds to one voice (part) indicated by colours of the lights. The music only progresses when the motion of the two are roughly in synchronisation.

\subsection{Construction}
\subsubsection{Keys}
The keys are laser-cut from aluminum of 5 mm thickness. We chose aluminum instead of an alloy since we would like a mellower timbre. Prior to production, we ordered a small number of samples from the fabricator to determine the acoustic characteristics of the material, i.e., timbre and the coefficient for resonant frequency. The resonant frequency of a vibrating metal bar is inversely proportional to the square root of its length. We thus calculate the desired lengths according to equal-temperament tuing, create the blueprints programmatically in OpenSCAD and send them for fabrication.

@@

Please refer to the attached video for a performance recording.

@@

The estimation algorithms in the pendulum enables it to act as a standalone sensor, adding to the toolkit of digital artistic expression. What about an electronic mimicry of a windbell? A big hanging simulated telescope? A sparkling earring?

>>>>

% Uncomment only one of the ones below
% \usepackage{anonymize} 		   %Uncomment this line to publish
\usepackage[blind]{anonymize}%Uncomment this line for blind review

@@

\label{key}
\numberofauthors{4}
\author{
\alignauthor
\anonymize{Lyu, S.} \\
  \affaddr{\anonymize{Academy of Arts \& Design, Tsinghua University}}\\
  \affaddr{\anonymize{Beijing, China}}\\
  \email{\anonymize{claire@ayu.land}}
\alignauthor
\anonymize{Li, H.} \\
  \affaddr{\anonymize{Academy of Arts \& Design, Tsinghua University}}\\
  \affaddr{\anonymize{Beijing, China}}\\
  \email{\anonymize{lihanxua22@mails.tsinghua.edu.cn}}
\and
\alignauthor
\anonymize{Wang, R.} \\
  \affaddr{\anonymize{Academy of Arts \& Design, Tsinghua University}}\\
  \affaddr{\anonymize{Beijing, China}}\\
  \email{\anonymize{wangrh22@mails.tsinghua.edu.cn}}
\alignauthor
\anonymize{Mi, H.}\titlenote{Corresponding author.}\\
  \affaddr{\anonymize{Academy of Arts \& Design, Tsinghua University}}\\
  \affaddr{\anonymize{Beijing, China}}\\
  \email{\anonymize{mhp@tsinghua.edu.cn}}
}
\date{6 February 2024}

@@

The instrument is accompanied by two Malletwand controllers and can operate in both solo and co-operative mode. Signals received from the controllers drive the passing of musical time in a predetermined musical piece, similar to a musical box. Co-operative mode is entered by two controllers being operated simultaneously, where each controller corresponds to one voice (part) indicated by colours of the lights. The music only progresses when the motion of the two are roughly in synchronisation.

\subsection{Construction}
\subsubsection{Keys}
The keys are laser-cut from aluminum of 5 mm thickness. We chose aluminum instead of an alloy since we would like a mellower timbre. Prior to production, we ordered a small number of samples from the fabricator to determine the acoustic characteristics of the material, i.e., timbre and the coefficient for resonant frequency. The resonant frequency of a vibrating metal bar is inversely proportional to the square root of its length. We thus calculate the desired lengths according to equal-temperament tuning, create the blueprints programmatically in OpenSCAD and send them for fabrication.

@@

Please refer to the attached video for a performance recording.~\footnote{The video is available at \url{http://bit.ly/42xa08Q} during review. We have removed traces of the authors' identities.}

@@

The estimation algorithms in the pendulum enable it to act as a standalone sensor, adding to the toolkit of digital artistic expression. What about an electronic mimicry of a windbell? A big hanging simulated telescope? A sparkling earring?
